Predicting the effects of character displacement on food-web dynamics

### 2019-03-11 manuscript_ECD_model.Rmd
add evolution results in McCann model and remove small points from all panels except for C (stability)

@@

where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$. $F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e. functional response) on resource $i$. 
This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g. zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g. seed size, see @Taper1985 for an example). 
In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.

To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$). 
At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a linear stability analysis [@Otto2007]. 
This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.

@@

@@

```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers can forage for both resources simultaneously."}
Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(a[ii]+a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + #, show.legend=FALSE) + size=5,
  geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (Fig. \ref{fig:McCann_Plot} in Appendix). This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (derivation in supplementary material):

@@

In the previous models, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$). In this more realistic model, however, whether the food web is locally stable depends on the parameters of both consumers and resources. Specifically, the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in supplementary material):

@@

Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig. \ref{fig:LS_Plot}A). Thus, character displacement always pushes the food web toward the boundary of local stability under this more realistic foraging scenario.

```{r McCann_Resources, include=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers exhibit a more realistic functional response."}
McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

McCann_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(a[ii]+a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Mac_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Stability_Mac_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + #size=5,
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  ylab(expression("Stability "(-lambda[max]))) + 
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  geom_hline(yintercept=0, linetype="dotted", color="grey") +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

LS_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

>>>>

*Number of words*: `r wordcountaddin::word_count("manuscript_ECD_model.Rmd")`

@@

where $r_i$ represents the intrinsic growth rate of resource $i$, $K_i$ represents the carrying capacity of resource $i$, $e_{ji}$ represents the conversion efficiency of resource $i$ into consumer $j$, and $m_j$ represents the mortality rate of consumer $j$. $F_{ji}(R_i)$ represents consumer $j$'s feeding rate (i.e. functional response) on resource $i$. 
This model is a useful characterization of a scenario where consumers compete for two distinct resources (e.g. zooplankton and benthic invertebrates) rather than if resources are better characterized by a continuous trait distribution (e.g. seed size, see @Taper1985 for an example). 
<!--
In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer. 
-->
To investigate the ecological consequences of divergent character displacement, I analyzed differences in resource abundances and food-web stability at equilibrium. 
An equilibrium is reached when the four rates of change in the above equations are 0, and solving the system at this point gives equilibrium abundances for each resource ($\hat R_1,\hat R_2$) and consumer ($\hat C_1,\hat C_2$). 
At the coexistence equilibrium (i.e., where $\hat R_1,\hat R_2, \hat C_1,\hat C_2 > 0$), I calculated resource abundances and performed a linear stability analysis [@Otto2007]. 
This stability analysis derives the dominant eigenvalue, $\lambda$, which determines whether (and how readily) the food web will return to equilibrium after a small perturbation.

@@

In the simplest scenario, consumers can forage for both resources simultaneously and their feeding rates increase linearly with resource abundance, such that $F_{ji}(R_i)=a_{ji}R_i$ where $a_{ji}$ is the attack rate of consumer $j$ on resource $i$ [@MacArthur1972]. 
The evolution of consumer attack rates in this model (and several extensions) have been analyzed in detail [@Lawlor1976; @Abrams1986], with the general result being divergent character displacement. 
I say consumers have undergone divergent character displacement if there evolved attack rates are more specialized, measured as $|\frac{a_{ii}}{a_{ii}+a_{ij}}-0.5|$, when evolving with a competing consumer.

@@

```{r MacArthur_Resources, include=FALSE, fig.cap="\\label{fig:MacArthur_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers can forage for both resources simultaneously."}
Mac_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(a[ii]+a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Mac_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
  #geom_point(data = filter(MacArthur_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

```{r LS_Resources, include=FALSE, fig.cap="\\label{fig:LS_Resources}Effect of character displacement on effective attack rates (A) and resource abundances (B) when consumers cannot forage for resources simultaneously."}
LS_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

LS_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color=factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

The evolutionary dynamics of consumer attack rates have not been examined in this model, but as with the other models [@LawlorSmith1976], I find that divergent character displacement is inevitable. 
This is because, in the absence of competition, consumer evolution is determined by both resource abundances and their habitat preferences, and reach an evolutionary stable strategy (ESS) when:

$$w_{ii}R_i^2\delta_{ii}=w_{ij}R_j^2\delta_{ij}$$ 
where $\delta_{ii}$ represents a small mutation in $a_{ii}$ and we assume a negative tradeoff in attack rates, such that $\delta_{ij}=-f(\delta_{ii})$. 
The square-terms for resource abundances indicate that they largely govern the evolutionary dynamics. 
This scenario favors the evolution of a generalist phenotype, as resource specialization will decrease that resource's abundance, thus favoring evolution toward the other resource until an ESS is reached.
In contrast, in the presence of a competitor, resource abundances will be equivalent at equilibrium (derivation in Appendix S2) and consumers reach an ESS when:

$$w_{ii}\delta_{ii}=w_{ij}\delta_{ij}$$

Assuming consumers have a preferred resource (i.e. $w_{ii}>0.5$), then consumers will only reach an ESS when $|\delta_{ii}|<|\delta_{ij}|$ (e.g. green line, Fig. \ref{MacArthur_Plot}A). 
Otherwise, they will evolve to be complete specialists on their preferred resource.

Divergent character displacement in this foraging scenario has the same general effect on food-web dynamics as the previous scenario (resources in different habitats, but linear functional response). 
<!--In terms of food-web dynamics, I observe the same general effect of ecological character displacement as before when resources occurred in distinct habitats, but consumers had a simpler functional response. -->
This is because resource abundances at equilibrium are governed by a similar dynamic (derivation in Appendix S2):
<!--
Even after adding more biological realism, I still observe the same qualitative relationship between character displacement and resource suppression (Fig. \ref{fig:McCann_Plot} in Appendix). This is because resource competition always leads to character divergence and equilibrium resource abundances are governed by similar dynamics  (derivation in supplementary material):
-->

@@

And since evolution favors consumer divergence onto their preferred resources, the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$) will always increase (Fig. \ref{fig:McCann_Plot}A), resulting in lower resource abundances (Fig. \ref{fig:McCann_Plot}B) and decreased food-web stability (Fig. \ref{fig:McCann_Plot}C).

In the previous models, character displacement influences food-web stability, but all of the food webs ultimately return to a stable equlibrium (because $\lambda_{max}<0$, derivation in Appendix S1 and S2). 
In this more realistic model, however, whether the food web is locally stable depends on the parameters of both consumers and resources. 
Using Routh-Hurwitz criteria, I found that the four-species food web will transition from having a locally stable equilibrium to a limit cycle under the following conditions (derivation in Appendix S2):

@@

<!--Previously, I showed that character displacement always increases the effective attack rate of consumers ($wa_{ii}+(1-w)a_{ij}$), regardless of the the shape of the tradeoff (Fig. \ref{fig:LS_Plot}A).--> 
Thus, character displacement always pushes the food web toward an unstable stucture in this more realistic foraging scenario.

```{r McCann_Resources, include=FALSE, fig.cap="\\label{fig:McCann_Resources}Effect of character displacement on total attack rates (A) and resource abundances (B) when consumers exhibit a more realistic functional response."}
McCann_attack <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

McCann_resources <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(McCann_displacements, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

Mac_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=aii+aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=a11+a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(a[ii]+a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Mac_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") +
  #geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

Stability_Mac_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), MacArthur_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=-1*max.Re.eigen, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=3, color="white", show.legend=FALSE) + 
  geom_point(data = filter(MacArthur_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=-1*max.Re.eigen, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  ylab(expression("Stability "(-lambda[max]))) + 
  scale_shape_manual(values=c(24,21))  +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  geom_hline(yintercept=0, linetype="dotted", color="grey") +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

LS_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

LS_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), LawlorSmith_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(LawlorSmith_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

@@

McCann_attack_asymm <- ggplot(tradeoff_df, aes(x=aii/(aii+aij), y=wii*aii+wij*aij, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=w11*a11+w12*a12, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(w*a[ii]+(1-w)*a[ij])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))

McCann_resources_asymm <- cbind.data.frame(select(tradeoff_df, n, aii, aij), McCann_2C_2R_tradeoff_dynamics) %>%
  ggplot(., aes(x=aii/(aii+aij), y=R1+R2, color = factor(n))) + 
  geom_line() +
  geom_point(data = filter(McCann_displacements_asymm, Time=="End"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=3, color="white") + 
  #geom_point(data = filter(McCann_displacements_asymm, Time=="Begin"), aes(x = a11/(a11+a12), y=R1+R2, fill=factor(n), shape=Competitor), size=1.5, color="white") +
  ylab(expression(R[i]+R[j])) +
  xlab(expression(frac(a[ii],a[ii]+a[ij]))) +
  scale_shape_manual(values=c(24,21)) +
  scale_color_manual(name="Tradeoff (n)", values=cbPalette)  +
  scale_fill_manual(name="Tradeoff (n)", values=cbPalette) +
  guides(fill = guide_legend(override.aes=list(shape=21)))
